Aspects of RDF Data Constraints in the Social, Behavioural and Economic Sciences

### 2015-04-30 paper/paper7/paper7.tex
intro

For research institutes, data libraries, and data archives,
RDF data validation according to predefined constraints is a much sought-after feature, 
particularly as this is taken for granted in the XML world.
Based on our work in the DCMI RDF Application Profiles Task Group and in cooperation with the W3C Data Shapes Working Group, we identified and published by today 82 types of constraints that are required by various stakeholders for data applications.
In this paper, we formulate 246 constraints of 53 different types on six different vocabularies (Disco, QB, SKOS, PHDD, DCAT, XKOS) and classify them according to the complexity level of their type and their severity level. For 114 of these constraints, we evaluate the data quality of 15,694 data sets (4.26 billion triples) of research data for the social, behavioural, and economic (SBE) sciences obtained from 33 SPARQL endpoints. 
Based on the results, we formulate several hypotheses to direct the further development of constraint languages.

@@

With the rise of the Web of Data, data professionals and institutions are very interested in having their data be discovered and used by publishing their data directly in RDF or at least publish accurate metadata about their data to facilitate data integration. Therefore, not only existing vocabularies like SKOS are used; 
recently, members of the \emph{SBE} and Linked Data community developed with the \emph{DDI-RDF Discovery Vocabulary (Disco)}\footnote{\url{http://rdf-vocabulary.ddialliance.org/discovery.html}} a means to expose \emph{DDI} metadata as Linked Data.

For constraint formulation and validation on RDF data, several languages exist, like \emph{Shape Expressions}, \emph{Resource Shapes} or \emph{Description Set Profiles}. \emph{OWL 2} is also used as a constraint language under a closed world assumption. With its direct support of validation via SPARQL, \emph{SPIN}%\footnote{\url{http://spinRDF.org/}}
is very popular and certainly plays an important role for future developments in this field.
%that more or less explicitly address the \emph{SBE} community.

In 2013, the W3C organized the RDF Validation Workshop\footnote{\url{http://www.w3.org/2012/12/rdf-val/}}, 
where experts from industry, government, and academia discussed first use cases for RDF constraint formulation and RDF data validation.
In 2014, two working groups on RDF validation have been established to develop a language to express constraints on RDF data: 
the W3C RDF Data Shapes working group\footnote{\url{http://www.w3.org/2014/rds/charter}} and the DCMI RDF Application Profiles task group\footnote{\url{http://wiki.dublincore.org/index.php/RDF-Application-Profiles}}.

Bosch and Eckert \cite{BoschEckert2014} collected the findings of these working groups and initiated a database of RDF validation requirements
which is available for contribution at \url{http://purl.org/net/rdf-validation}.
The intention is to collaboratively collect case studies, use cases, requirements, and solutions regarding RDF validation in a comprehensive and structured way. 
The requirements are classified to better evaluate existing solutions and each requirement is directly mapped to a constraint type which is expressible by at least one constraint language.
As \emph{SPIN} is the within the RDF validation community accepted and widely adopted way to check constraints on RDF data, Bosch and Eckert \cite{BoschEckert2014-2} use \emph{SPIN} as basis to define a
validation environment (available at \url{http://purl.org/net/rdfval-demo}) in which the validation of any constraint language\footnote{The only limitation is that constraint languages must be represented in RDF} can be implemented by representing them in SPARQL. 
The SPIN engine checks for each resource if it satisfies all constraints, which are associated with its assigned classes, and generates a result RDF graph containing information about all constraint violations.

%This paper aims to address two main \textbf{audiences}: 
%(1) metadata practitioners seeking for how to represent metadata and data on different aggregation levels in RDF and
%(2) metadata providers and harvesters ensuring high quality data and metadata by validating RDF data according to certain constraints.

@@

In this paper, we discuss constraints on RDF data in general. Note that the data represented in RDF can be data in the sense of SBE sciences, but also metadata about published or unpublished data. We generally refer to both simply as RDF data and only distinguish between data and metadata in the data set descriptions and in the case that it matters for the purpose of this paper.

@@

%\section{Running Example}
%\label{running-example}
\section{Common Vocabularies in SBE Sciences}
\label{rdf-representation}
\tb{Ich habe XKOS und DCAT noch ganz am Ende eingeführt (jeweils nur 1 Satz)}
%In this section, we describe how data in rectangular format (\emph{PHDD}) and metadata on person-level data sets (\emph{Disco}), aggregated data sets (\emph{QB}), thesauri (\emph{SKOS}), and statistical classifications (\emph{XKOS}) are represented in RDF\footnote{The complete running example in RDF is available at: \url{https://github.com/boschthomas/rdf-validation/tree/master/data/running-example}} and how therefore reused vocabularies are interrelated.

For our evaluation, we examined six different vocabularies commonly used in or developed for the SBE sciences which are briefly introduced in the following. For three of them, we analysed actual data according to constraint violations. For PHDD, there is not yet enough data available so we limit our examination to the vocabulary itself.

@@

%\textbf{Organizations, Hierarchies, and Classifications.}
%\textbf{Reusing SKOS.}
The \emph{Simple Knowledge Organization System (SKOS)} is reused multiple times to build \emph{SBE} vocabularies.
%Variables are constructed out of values and/or (un)ordered code lists.
The codes of the variable \emph{Education at pre-school} (number of education hours per week) are modeled as \emph{skos:Concepts} and 
a \emph{skos:OrderedCollection} organizes them in a particular order within a \emph{skos:memberList}.
A variable may be associated with a theoretical concept (\emph{skos:Concept}). 
\emph{skos:narrower} builds the hierarchy of theoretical concepts within a \emph{skos:ConceptScheme} of a series.
The variable \emph{Education at pre-school} is assigned to the theoretical concept \emph{Child Care} which is the narrower concept of \emph{Education} - one of the top concepts of the series \emph{EU-SILC}.
Controlled vocabularies (\emph{skos:ConceptScheme}), serving as extension and reuse mechanism,
organize types (\emph{skos:Concept}) of descriptive statistics (\emph{disco:SummaryStatistics}) like minimum, maximum, and arithmetic mean.
\emph{XKOS}\footnote{\url{https://github.com/linked-statistics/xkos}} is a SKOS extension to describe formal statistical classifications like the International Standard Classification of Occupations (\emph{ISCO}). 
and the Statistical Classification of Economic Activities in the European Community \emph{NACE}.
\emph{DCAT} enables to represent data sets inside of data collections like portals, repositories, catalogs, and archives
which serve as typical entry points when searching for data.

%\textbf{Searching for (Meta)data.}
%\emph{DCAT} enables to represent data sets inside of data collections like portals, repositories, catalogs, and archives
%which serve as typical entry points when searching for data.
%Users search for aggregated and person-level data records (\emph{dcat:CatalogRecord}) inside data catalogs (\emph{dcat:Catalog}). 
%As search differs depending on the users’ information need,
%users may only search for records' metadata (e.g., \emph{dcterms:title}, \emph{dcterms:description}), 
%or may formulate more sophisticated queries on aggregated and person-level data sets (\emph{dcat:Dataset}) or their
%distributions (\emph{dcat:Distribution}) which are part of the records. 
%Often, users search for data sets covering particular topics (\emph{dcat:keyword}, \emph{dcat:theme}), time periods (\emph{dcterms:temporal}),  or  locations (\emph{dcterms:spatial}), 
%or for certain formats in which the data distribution is available (\emph{dcterms:format}).

@@

	%\item First, we assigned each constraint type to exactly one of the disjoint sets of constraint types 
%in order to get an overview how many constraint types in relation to the total amount of constraint types are extractable from vocabularies ($\mathcal{CT}_{B}$), are easily definable ($\mathcal{CT}_{S}$), and are rather difficult to specify ($\mathcal{CT}_{C}$). 
In close collaboration with several \emph{SBE} domain experts, we defined 246 constraint 
of 53 different types on six vocabularies (\emph{Disco}, \emph{QB}, \emph{SKOS}, \emph{PHDD}, \emph{DCAT}, \emph{XKOS}) and classified them according to the complexity level of their type and their severity level. 
For 114 of these constraints, we evaluated the data quality of 15,694 data sets (4.26 billion triples) of \emph{SBE} research data on three common vocabularies in \emph{SBE} sciences (\emph{Disco}, \emph{QB}, \emph{SKOS}) obtained from 33 SPARQL endpoints.
We distinct two classes of vocabularies: 
(1) well-established vocabularies (e.g., \emph{QB}, \emph{SKOS}) which are widely adopted and accepted and 
(2) newly developed vocabularies (e.g., \emph{Disco} which will be published in 2015) which are either recently published or are still in the publication process.
%It is likely that such a vocabulary is still subject of constant change, that published data may not be consistent with its current version,
%and that early adopters did not properly understand its formal specification.
%\ke{ergebnisse, dann hypothesen, dann kritische auseinandersetzung}

@@

As the evaluation is based on three vocabularies, 
we cannot make valid general statements for all vocabularies,
but we can formulate several hypotheses to direct the further development of constraint languages.
As these hypotheses cannot be proved yet, 
they still have to be verified or falsified
by evaluating the quality of data represented by further well-established and newly developed vocabularies.
%More than the half of them are $\mathcal{CT}_{B}$, nearly a third $\mathcal{CT}_{S}$, and only a sixth $\mathcal{CT}_{C}$ constraint types.
%For \emph{Disco}, \emph{QB}, and \emph{SKOS}, more than 50\% of the instantiated constraint types are $\mathcal{CT}_{B}$ constraint types 
%(for \emph{QB} even three quarters).
%\emph{Existential quantifications} (\emph{R-86}, 32.4\%, \emph{Disco}), \emph{data model consistency} (31.4\%, \emph{QB}), and \emph{structure} (28.6\%, \emph{SKOS}) are the constraint types the most constraints are instantiated from.
%\tb{1. Fakten (diese Fakten bewirken eine Erhöhung der Datenqualität) 2. Hypothese (unsere Interpretation der Ergebnisse, unsere Behauptung)}
Almost 1/2 of all 213 constraints and more than 50\% of the \emph{Disco} and the \emph{QB} constraints are \emph{vocabulary constraints}.
The \emph{SKOS} constraints are nearly to the same extend \emph{vocabulary}, \emph{simple}, and \emph{complex constraints}. 
%Hence, a significant amount of 46 \% of the constraints are directly extractable from formal specifications of well-established and newly defined vocabularies. 
\begin{hyp}
As a significant amount of 46 \% of the constraints are vocabulary constraints which can be expressed by modeling languages like RDF, RDFS, and OWL 2,
the further development of constraint languages should concentrate on expressing simple and especially complex constraints which up to now in most cases can only be expressed by plain SPARQL.   
\end{hyp}

Nearly 1/2 of all violations are caused by \emph{complex constraints}, 1/3 by \emph{simple constraints}, and 1/5 by \emph{vocabulary constraints}.
The fact that only 1/5 of all violations result from \emph{vocabulary constraints}, 
even though, 46\% of all constraints are \emph{vocabulary constraints},
indicates good data quality for all vocabularies with regard to their formal specifications.
\begin{hyp}
For all vocabularies, data corresponds to their formal specifications which demonstrates that constraint formulation in general works. 
\end{hyp}

2/3 of the \emph{Disco} violations result from \emph{vocabulary constraints},
\emph{QB} violations are almost only raised by \emph{complex constraints}, and 
nearly 80\% of the \emph{SKOS} violations are caused by \emph{simple constraints}.
For well-established vocabularies, \emph{vocabulary constraints} are almost completely satisfied\footnote{e.g. only 1.777 \emph{QB} violations} which indicates good data quality according to formal specifications of vocabularies.
%This demonstrates that data providers take account of the formal specifications of well-established vocabularies.
For newly defined vocabularies, however, 2/3 of all violations are raised by \emph{vocabulary constraints}
which indicates bad data quality with regard to the formal specifications of vocabularies.

\begin{hyp}
Data represented by well-established vocabularies corresponds to formal specifications of these vocabularies which demonstrates that constraint formulation in general works. 
Data represented by newly developed vocabularies, in contrast, does not correspond to formal specifications of these vocabularies.
\end{hyp}

It is likely that a newly developed vocabulary is still subject of constant change
and that early adopters did not properly understand its formal specification.
Thus, published data may not be consistent with its current draft or version.
In case newly developed vocabularies turn into well-established ones,
data providers are experienced in publishing their data in conformance with these vocabularies, 
and the formal specifications are more elaborated and therefore clearer. 
As a consequence, vocabulary constraints are satisfied which leads to better data quality.

\begin{hyp}
A significant amount of 47\% of the violations refer to complex constraints that are not easily, concisely, and intuitively expressible in existing constraint languages which confirms the necessity to provide suitable constraint languages.
\end{hyp} 
In general, \emph{vocabulary} and \emph{simple constraints} are easily, concisely, and intuitively expressible by either modeling languages (e.g., \emph{RDF}, \emph{RDFS}, \emph{OWL 2}) or constraint languages (e.g., \emph{ShEx}, \emph{ReSh}, \emph{SPIN}). 
This is not the case, however, for \emph{complex constraints} which in most cases are still only expressible by plain \emph{SPARQL}. 
As almost 1/2 of all violations are caused by \emph{complex constraints}, 
data quality can be significantly improved when suitable constraint languages are developed which enable to define \emph{complex constraints} in an easy, concise, and intuitive way.
%Therefore, existing constraint languages have to be extended or new constraint languages have to be developed.

@@

1/2 of all constraints are \emph{error constraints} and
10\% of all constraints are \emph{warning constraints}.
\emph{Informational constraints} caused 1/3 and
\emph{warning constraints} 2/3 of all violations.
As the percentage of severe violations is very low for all vocabularies,
data quality is high with regard to the severity level of constraints.

\begin{hyp}
The percentage of severe violations is very low, compared to about 2/3 of warning violations and 1/3 of informational violations, which implies that proper constraint languages can significantly improve the data quality beyond fundamental requirements.
\end{hyp}

>>>>

For research institutes, data libraries, and data archives,
RDF data validation according to predefined constraints is a much sought-after feature, 
particularly as this is taken for granted in the XML world.
Based on our work in the DCMI RDF Application Profiles Task Group and in cooperation with the W3C Data Shapes Working Group, we identified and published by today 82 types of constraints that are required by various stakeholders for data applications.
In this paper, we formulate 246 constraints of 53 different types on six different vocabularies (Disco, QB, SKOS, PHDD, DCAT, XKOS) and classify them according to the complexity level of their type and their severity level. For 114 of these constraints, we evaluate the data quality of 15,694 data sets (4,26 billion triples) of research data for the social, behavioural, and economic (SBE) sciences obtained from 33 SPARQL endpoints. \ke{check numbers}
Based on the results, we formulate several hypotheses to direct the further development of constraint langugaes.

@@

With the rise of the Web of Data, data professionals and institutions are very interested in having their data be discovered and used by publishing their data directly in RDF or at least publish accurate metadata about their data to facilitate data integration. Therefore, not only established vocabularies like SKOS are used; 
recently, members of the \emph{SBE} and Linked Data community developed with the \emph{DDI-RDF Discovery Vocabulary (Disco)}\footnote{\url{http://rdf-vocabulary.ddialliance.org/discovery.html}} a means to expose \emph{DDI} metadata as Linked Data.

For constraint formulation and validation on RDF data, several languages exist or are currently developed, like \emph{Shape Expressions}, \emph{Resource Shapes} or \emph{Description Set Profiles}. \emph{OWL 2} is also used as a constraint language under a closed world assumption. With its direct support of validation via SPARQL, \emph{SPIN}%\footnote{\url{http://spinRDF.org/}}
is very popular and certainly plays an important role for future developments in this field. It is particularly interesting as a means to validate arbitrary constraint languages by mapping them to SPARQL \cite{BoschEckert2014-2}. Yet, there is no clear favorite and none of the languages is able to meet all requirements raised by data practitioners. Further research and development therefore is needed.

In 2013, the W3C organized the RDF Validation Workshop\footnote{\url{http://www.w3.org/2012/12/rdf-val/}}, 
where experts from industry, government, and academia discussed first use cases for RDF constraint formulation and RDF data validation.
In 2014, two working groups on RDF validation have been established to develop a language to express constraints on RDF data: 
the W3C RDF Data Shapes working group\footnote{\url{http://www.w3.org/2014/rds/charter}} and the DCMI RDF Application Profiles task group\footnote{\url{http://wiki.dublincore.org/index.php/RDF-Application-Profiles}} which among others bundles the requirements of data institutions of the cultural heritage and SBE sector and represents them in the W3C group.

Within the DCMI task group, a collaboratively curated database of RDF validation requirements has been created which contains the findings of the working groups based on various case studies provided by data institutions \cite{BoschEckert2014}. It is publicly available and open for further contributions\footnote{Online at \url{http://purl.org/net/rdf-validation}}.

The database connects requirements to use cases, case studies and implementations and forms the basis of this paper. To gain a better understanding about the role of certain requirements for data quality and in order to direct the further development of constraint langugaes, we collected constraints for commonly used vocabularies in the SBE domain, either from the vocabularies themselves or from domain and data experts. All in all, this lead to TODO constraints on TODO vocabularies. We let the experts classify the constraints according to the severity if they are violated. Furthermore, we classified the type of each constraint (corresponding to a requirement) based on its complexity ranging from types commonly found in vocabulary specificatons (e.g., domain and range), over types that are simply stated using common constraint languages (e.g., cardinality restrictions) to complex types that involve complex data structures or need more sophisticated langugaes to be easily expressible (e.g., constraints on network topologies).\ke{Thomas: Ok so?}

As we do not want to base our conclusions on the evaluations of vocabularies and constraint definitions alone, we conducted a large-scale experiment and evaluated the data quality of 15,694 data sets (4,26 billion triples) of research data for the social, behavioural, and economic (SBE) sciences obtained from 33 SPARQL endpoints.

@@

In this paper, we discuss constraints on RDF data in general. Note that the data represented in RDF can be data in the sense of SBE sciences, but also metadata about published or unpublished data. We generally refer to both simply as RDF data and only distinguish between data and metadata in the dataset descriptions and in the case that it matters for the purpose of this paper.\tb{hab ich eingefügt, ist der absatz von dir...}

@@

%\section{Running Example}
%\label{running-example}
\section{Common Vocabularies in SBE Sciences}
\label{rdf-representation}
\tb{jeweils ein satz zu DCAT und XKOS}
%In this section, we describe how data in rectangular format (\emph{PHDD}) and metadata on person-level data sets (\emph{Disco}), aggregated data sets (\emph{QB}), thesauri (\emph{SKOS}), and statistical classifications (\emph{XKOS}) are represented in RDF\footnote{The complete running example in RDF is available at: \url{https://github.com/boschthomas/rdf-validation/tree/master/data/running-example}} and how therefore reused vocabularies are interrelated.

For our evaluation, we examined four different vocabularies commonly used in or developed for the SBE sciences which are briefly introduced in the following. For three of them, we analysed actual data according to constraint violations. For PHDD, there is not yet enough data available so we limit our examination to the vocabulary itself.\ke{Neuer Absatz von mir...}

@@

%\textbf{Organizations, Hierarchies, and Classifications.}
%\textbf{Reusing SKOS.}
The \emph{Simple Knowledge Organization System (SKOS)} is reused multiple times to build \emph{SBE} vocabularies.
%Variables are constructed out of values and/or (un)ordered code lists.
The codes of the variable \emph{Education at pre-school} (number of education hours per week) are modeled as \emph{skos:Concepts} and 
a \emph{skos:OrderedCollection} organizes them in a particular order within a \emph{skos:memberList}.
A variable may be associated with a theoretical concept (\emph{skos:Concept}). 
\emph{skos:narrower} builds the hierarchy of theoretical concepts within a \emph{skos:ConceptScheme} of a series.
The variable \emph{Education at pre-school} is assigned to the theoretical concept \emph{Child Care} which is the narrower concept of \emph{Education} - one of the top concepts of the series \emph{EU-SILC}.
Controlled vocabularies (\emph{skos:ConceptScheme}), serving as extension and reuse mechanism,
organize types (\emph{skos:Concept}) of descriptive statistics (\emph{disco:SummaryStatistics}) like minimum, maximum, and arithmetic mean.
%\emph{XKOS}\footnote{\url{https://github.com/linked-statistics/xkos}} is a SKOS extension to describe formal statistical classifications like the International Standard Classification of Occupations (\emph{ISCO}). 
%and the Statistical Classification of Economic Activities in the European Community \emph{NACE}.

%\textbf{Searching for (Meta)data.}
%\emph{DCAT} enables to represent aggregated and person-level data sets inside of data collections like portals, repositories, catalogs, and archives
%which serve as typical entry points when searching for data.
%Users search for aggregated and person-level data records (\emph{dcat:CatalogRecord}) inside data catalogs (\emph{dcat:Catalog}). 
%As search differs depending on the users’ information need,
%users may only search for records' metadata (e.g., \emph{dcterms:title}, \emph{dcterms:description}), 
%or may formulate more sophisticated queries on aggregated and person-level data sets (\emph{dcat:Dataset}) or their
%distributions (\emph{dcat:Distribution}) which are part of the records. 
%Often, users search for data sets covering particular topics (\emph{dcat:keyword}, \emph{dcat:theme}), time periods (\emph{dcterms:temporal}),  or  locations (\emph{dcterms:spatial}), 
%or for certain formats in which the data distribution is available (\emph{dcterms:format}).

@@

	%\item First, we assigned each constraint type to exactly one of the disjoint sets of constraint types 
%in order to get an overview how many constraint types in relation to the total amount of constraint types are extractable from vocabularies ($\mathcal{CT}_{B}$), are easily definable ($\mathcal{CT}_{S}$), and are rather difficult to specify ($\mathcal{CT}_{C}$). 
In close collaboration with several \emph{SBE} domain experts, we defined 246 constraint 
of 53 different types on six vocabularies (\emph{Disco}, \emph{QB}, \emph{SKOS}, \emph{PHDD}, \emph{DCAT}, \emph{XKOS}) and classified them according to the complexity level of their type and their severity level. 
\ke{Alle vokabulare in section "vocabularies" eingeführt?}\tb{mache ich}
For 114 of these constraints, we evaluated the data quality of 15,694 data sets (4,26 billion triples) of \emph{SBE} research data on three common vocabularies in \emph{SBE} sciences (\emph{Disco}, \emph{QB}, \emph{SKOS}) obtained from 33 SPARQL endpoints.
We distinct two classes of vocabularies: 
(1) well-established vocabularies (e.g., \emph{QB}, \emph{SKOS}) which are widely adopted and accepted and 
(2) newly developed vocabularies (e.g., \emph{Disco} which will be published in 2015) which are either recently published or are still in the publication process.
%It is likely that such a vocabulary is still subject of constant change, that published data may not be consistent with its current version,
%and that early adopters did not properly understand its formal specification.
Based on this evaluation we formulate hypotheses.
Nevertheless, we have to be careful making general statements for all vocabularies 
as these hypotheses still have to be verified or falsified in terms of future work 
by evaluating the quality of data represented by additional well-established and newly developed vocabularies.
\ke{das ist zu früh finde um schon einzuschränken, erst mal ergebnisse, dann hypothesen, dann kritische auseinandersetzung}

@@

%More than the half of them are $\mathcal{CT}_{B}$, nearly a third $\mathcal{CT}_{S}$, and only a sixth $\mathcal{CT}_{C}$ constraint types.
%For \emph{Disco}, \emph{QB}, and \emph{SKOS}, more than 50\% of the instantiated constraint types are $\mathcal{CT}_{B}$ constraint types 
%(for \emph{QB} even three quarters).
%\emph{Existential quantifications} (\emph{R-86}, 32.4\%, \emph{Disco}), \emph{data model consistency} (31.4\%, \emph{QB}), and \emph{structure} (28.6\%, \emph{SKOS}) are the constraint types the most constraints are instantiated from.

\tb{Ich würde zu Beginn der Auswertung, nach den Tabellen, erklären, dass wir anhand der drei Vokabulare keine gesicherten allgemeinen Aussagen  machen können, aber dass wir im folgenden verschiedene Hypthesen aufstellen, die auf die Auswertung gestützt sind. / ann also erst die Fakten berichten, wie zum Beispiel bestimmte Prozentwerte / Und dann die Hypothese als Schluss, z..B. dass die Datenqualität bezüglich in den etablierten Vokabularen hinterlegter Constraints sehr gut ist, was darauf hindeutet, dass das hinterlegen von constraints sinnvoll ist und von den datenanbietern auch beachtet wird.
ganz eigentlich ist ja sogar nur das zweite unsere hypothese
die datenqualität in unserem sample IST gut, das ist auch keine hypothese / es muss halt klar werden, dass wir die hypothese im anschluss nicht beweisen, sondern dass wir sie lediglich postulieren, allerdings nicht komplett aus der luft sondern als interpretation unserer ergebnisse / Was zwar kein beweis ist wegen geringer zahl der vokabulare, aber eben auch nicht nix}

Almost 1/2 of all 213 constraints and more than 50\% of the \emph{Disco} and the \emph{QB} constraints are \emph{vocabulary constraints}.
The \emph{SKOS} constraints are equally distributed among the three disjoint sets of constraint types. 
\begin{hyp}
A significant amount of 46 \% of the constraints are directly extractable from formal specifications of well-established and newly defined vocabularies. 
\end{hyp}

Nearly 1/2 of the violations are caused by \emph{complex constraints}, 1/3 by \emph{simple constraints}, and 1/5 by \emph{vocabulary constraints}.

\begin{hyp}
The fact that only 1/5 of all violations result from vocabulary constraints, 
even though, 46\% of all constraints are vocabulary constraints,
indicates good data quality in general (according to formal specifications of vocabularies).
\end{hyp}

2/3 of the \emph{Disco} violations result from \emph{vocabulary constraints},
almost only \emph{complex constraints} raised \emph{QB} violations, and 
nearly 80\% of the \emph{SKOS} violations are caused by \emph{simple constraints}.

\begin{hyp}
For well-established vocabularies, vocabulary constraints are almost completely satisfied\footnote{e.g. only 1.777 \emph{QB} violations} which indicates good data quality (according to formal specifications of vocabularies) and which demonstrates that constraint formulation in general works.
For newly defined vocabularies, however, 2/3 of all violations are raised by vocabulary constraints
which indicates good data quality (according to formal specifications of vocabularies).
\end{hyp}

It is likely that a newly developed vocabulary is still subject of constant change, 
that published data may not be consistent with its current version,
and that early adopters did not properly understand its formal specification.

\begin{hyp}
A significant amount of 47\% of the violations refer to complex constraints that are not easily expressible in existing languages which confirms the necessity to provide suitable constraint languages.
\end{hyp} 
\emph{Vocabulary} and \emph{simple constraints} can be expressed easily, concisely, and intuitively by either modeling languages (e.g., \emph{RDF}, \emph{RDFS}, \emph{OWL 2}) or constraint languages (e.g., \emph{ShEx}, \emph{ReSh}, \emph{SPIN}). 
This is not the case, however, for \emph{complex constraints} which in most cases still can only be expressed by plain \emph{SPARQL}. 
As almost 1/2 of all violations are caused by \emph{complex constraints}, 
data quality can be significantly improved when \emph{complex constraints} can also be expressed more easily, concisely, and intuitively.
Therefore, existing constraint languages have to be extended or new constraint languages have to be developed.

63\% of all violations are caused by \emph{warning constraints} which are only 10\% of all constraints. 
Only 6\% of the violations are caused by \emph{error constraints} which are 49\% of all constraints.

\begin{hyp}
In general, data quality is high (according to severity levels) since only 6\% of the violations are caused by error constraints which are 49\% of all defined constraints.
\end{hyp}

\begin{hyp}
The percentage of severe violations is very low (6\%), compared to about 2/3 of warning violations and 1/3 of informational violations, which implies that proper constraint languages can significantly improve the data quality beyond fundamental requirements.
\end{hyp}

@@
